Physcraper: A Python package for continually updated phylogenetic trees using the Open Tree of Life

### 2021-04-06 2.4-discussion-ideas.Rmd
tidying up

>>>>

There is ample evidence that increased taxon sampling benefits phylogenetic inference. It breaks long branches (Hillis 1996), increases phylogenetic accuracy (Zwickl and Hillis, 2002), resolves phylogenetic conflict between gene trees (Hedtke et al. 2006), and resolves contentious phylogenetic relationships (Townsend & Lopez-Giraldez 2010; Natsidis et al. 2019), reveal hidden orthologs and paralogs (Martin-Duran et al. 2017). It has been the focus to give insight into phylogenetic relationships of many different taxa.
Applied phylogenetic analyses have largely focused on improving taxon sampling to gain better insight in a group’s evolutionary history:
in fish (Salmonidae, Crête et al. 2012; flounders, Vinnikov et al. 2018),
birds (Wang et al. 2017)
Gastropods (Oskars et al. 2015)
Insects (Hemiptera, Song et al. 2019; Diptera, Shin et al. 2018; Lepidopetra, Wilson 2011; Orthoptera, Vandergast et al. 2017),
amphibians (Peloso et al. 2016)
Mammals (Cetartiodactyla, Agnarsson & May-Collado 2008)
Squamate reptiles (Ketchum & Benson 2010)
turtles (Vitek et al. 2018)
flowering plants (Cornales, Xiang et al. 2011; Caryophyllales, Crawley & Hilu 2012; Campanula, Mansion et al. 2012, Rosids, Sun et al. 2020)
unicellular organisms (Ciliates, Tang & Zhao 2016; Xu et al. 2021; chlamydomonas, Nakada et al. 2019)
Spiders (Kallal et al. 2020)
Sponges (Alvizu et al. 2018)
fungi (Prsanna et al. 2020)
biogeography (bees, Kayaalp et al. 2017; gallus, )
Increased taxon sampling also improves time of divergence estimates (Schulte 2013; Soares & Schrago 2015)

### 2021-04-06 sources/bmc_template_submission/bmc_article.tex
tidying up

An online literature review in June 2020 (Google scholar search for "ilex phylogeny")
reveals that there are several published phylogenies showing relationships within
the hollies \cite{cuenoud2000molecular, manen2010history, setoguchi2000intersectional,
selbach2009new}, but only two have data publicly available \cite{gottlieb2005molecular, yao2020phylogeny}.
\cite{gottlieb2005molecular} made original tree and alignment available in TreeBASE
(study 1091 \cite{treebase1091}). This tree (Gottlieb2005 tree from now on) samples 41 species, is available from
OpenTree's Phylesystem (study pg\_2827 \cite{pg2827}),
and has been integrated into OpenTree's synthetic tree \cite{mrcaott68451ott89474}.
The most recent \textit{Ilex} tree \cite{yao2020phylogeny} (Yao2020 tree from now on) is available in
OpenTree's Phylesystem (study ot\_1984 \cite{ot1984}),
and in the DRYAD repository \cite{yao2020dryad}.
With 175 tips, the Yao2020 tree \cite{yao2020phylogeny} is the best sampled phylogeny available for
the genus \textit{Ilex}.

@@

Physcraper generates gene trees, which individually do not capture the full complexity
of species' evolutionary history \cite{song2012resolving}.
However, Physcraper facilitates
gathering alignments and gene trees for multiple loci from a group of interest,
that together can be used to reconstruct species trees taking into account coalescent
processes with ASTRAL \cite{mirarab2014astral} or SVD Quartets \cite{chifman2014quartet}).
Rigorous analyses of multiple loci allows for more complex evolutionary models
than analyses of large genomic data sets, and can provide better evolutionary estimates.

>>>>

An online literature review in June 2020 (Google scholar search for "ilex phylogeny")
reveals that there are several published phylogenies showing relationships within
the hollies \cite{cuenoud2000molecular, manen2010history, setoguchi2000intersectional,
selbach2009new}, but only two have data publicly available \cite{gottlieb2005molecular, yao2020phylogeny}.
\cite{gottlieb2005molecular} made original tree and alignment available in TreeBASE
(study 1091 \cite{treebase1091}). This tree (Gottlieb2005 tree from now on) samples 41 species, is available from
OpenTree's Phylesystem (study pg\_2827 \cite{pg2827}),
and has been integrated into OpenTree's synthetic tree \cite{mrcaott68451ott89474}.
The most recent \textit{Ilex} tree \cite{yao2020phylogeny} (Yao2020 tree from now on) is available in
OpenTree's Phylesystem (study ot\_1984 \cite{ot1984}),
and in the DRYAD repository \cite{yao2020dryad}.
With 175 tips, the Yao2020 tree \cite{yao2020phylogeny} is the best sampled phylogeny available for
the genus \textit{Ilex}.
In order to showcase Physcraper's performance, we updated
the tree with the smallest number of tips (the Gottlieb2005 tree) with the expectation
to use the tree with the largest number of tips (the Yao2020 tree) as a standard
for comparison and verification of Physcraper's results.

@@

Physcraper generates gene trees, which individually do not capture the full complexity
of species' evolutionary history \cite{song2012resolving}.
However, Physcraper facilitates
gathering alignments and gene trees for multiple loci from a group of interest,
that together can be used to reconstruct species trees taking into account coalescent
processes with ASTRAL \cite{mirarab2014astral} or SVD Quartets \cite{chifman2014quartet}.
Rigorous analyses of multiple loci allows for more complex evolutionary models
than analyses of large genomic data sets, and can provide better evolutionary estimates.
